Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-09 paper/galsize_paper.tex
Updated discussion of relation to previous work

Numerous previous analyses have employed the \citet{yang_etal05b} group catalog to ``directly" study the relationship between galaxy mass, size, and environment. For example, by treating observed groups as genuine dark matter halos, several analyses have found that the $\mstar-\rhalf$ relation of early-type galaxies exhibits weak, if any, environmental dependence \citep{weinmann_etal08,huertas_company_etal13b,shankar_etal14}. Meanwhile, \citet{weinmann_etal08} additionally find that for late-type galaxies, there is a strong correlation between disk size and central/satellite designation. Qualitatively, our clustering measurements shown in Figure \ref{fig:colorclustering} are in good agreement with these findings: once galaxies have been color-selected, only the clustering of blue galaxies shows significant dependence upon $\rhalf.$ Our clustering results are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass.

\textcolor{blue}{AK: when I was reading this paragraph, I had trouble digesting its meaning... Also it definitely broke the flow of the narrative for me.}
As discussed in \citet{spindler_wake17}, it is entirely possible that size differences between centrals and satellites of the same mass can be accounted for by mutual covariance with an additional variable such as star formation rate or morphological type \citep[see also][for an explicit demonstration of this scenario]{lilly_carollo16}. In particular, suppose that centrals and satellites of the same mass have different early-type fractions as indicated in \citet{weinmann_etal06}, and that early- and late-type galaxies exhibit universal, but distinct, $\mstar-\rhalf$ relations. In such a case, centrals would be larger than satellites of the same mass, even though {\em early-type centrals} would have the same sizes as {\em early-type satellites}. This scenario could also explain the sign-change and weakening of $\rhalf-$dependent clustering measurements in the present work.

@@

The scope of this work is to serve as a pilot study in which we identify the chief ingredients that can influence the $\rhalf-$dependence of galaxy clustering. While our results suggest that sizes of galaxies are linearly correlated with the
virial radius of halos at the moment the halo mass reached its peak and that sizes of satellite galaxies were set {\em prior} to infall, our model is probably does not capture galaxy properties and their trends in full quantitative detail.
For example, the clustering ratios of this model shown in Figure \ref{fig:clustering_ratio_upshot} do capture the overall trends, but also don't match the observed statistics fully.
 For present purposes, we have chosen to focus on exploring bracketing cases of qualitative ingredients, rather than fine-tuning the model.

Since the clustering signal is strongly influenced by differences between centrals and satellites, then the satellite fraction $F_{\rm sat}(\mstar)$ plays an important role in $\rhalf-$dependent clustering. For {\em fixed} scaling relations $\median{\rhalf^{\rm cens}}{\mhalo}$ and $\median{\rhalf^{\rm sats}}{\mhalo},$ models with different satellite fractions will exhibit different $\rhalf-$clustering ratios because $F_{\rm sat}(\mstar)$ controls the relative weighting of the two populations. An explicit demonstration of this point appears in the Appendix, which repeats the analyses in the main body of the text, but for a subhalo catalog with no orphan correction.

@@

Figure \ref{fig:censatsizeratios} shows that the satellites in $\mhalo\approx10^{12}\msun$ halos can be as large as $\lesssim40-50\%$ smaller than their central galaxy counterparts. Since the satellite fraction varies quite significantly with morphology and/or color, we have demonstrated that it is entirely plausible that most of the observed environmental trends of galaxy size can be understood simply in terms of central vs. satellite size differences. The latter reflects not evolution of galaxies inside host halos, but is due to the fact that satellites have reached their peak mass at higher redshifts than the centrals. While it remains to be seen whether this explanation ends up comprising the whole story, one thing is now clear: it is not possible to reliably analyze environmental trends of galaxy $\rhalf$ without properly accounting for satellites. %Because of the large systematics that are common in contemporary group-finding methods \citep{campbell_etal15}, we consider forward modeling methods to be a requisite for continued progress on this topic.

>>>>

Numerous previous analyses have employed the \citet{yang_etal05b} group catalog to ``directly" study the relationship between galaxy mass, size, and environment. For example, by treating observed groups as genuine dark matter halos, several analyses have found that the $\mstar-\rhalf$ relation of early-type galaxies exhibits weak, if any, environmental dependence \citep{weinmann_etal08,huertas_company_etal13b,shankar_etal14}.

As shown in \citet{campbell_etal15}, group-finding methods are subject to significant systematics, particularly when analyzing trends based on color: contemporary algorithms {\em generically} wash out differences between centrals and satellites; trends can even be artificially created that are not truly present. Our clustering-based, forward modeling methods are immune to such systematics, and so it is encouraging that our findings are commensurable with these results. Once galaxies have been color-selected, Figure \ref{fig:colorclustering} shows that the observed clustering trends nearly vanish in magnitude, and reverse sign. Insofar as blue galaxies are largely disk-dominated, the clustering results in Figure \ref{fig:colorclustering} are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass.

@@

The scope of this work is to serve as a pilot study in which we identify the chief ingredients that can influence the $\rhalf-$dependence of galaxy clustering. While our results suggest that sizes of galaxies are linearly correlated with the
virial radius of halos at the moment the halo mass reached its peak, and also that sizes of satellite galaxies were set {\em prior} to infall, our model does not capture galaxy properties and their trends in full quantitative detail.
For example, the clustering ratios of this model shown in Figure \ref{fig:clustering_ratio_upshot} do capture the overall trends, and also do not match the observed clustering statistics at the level of $\chi^2.$ For present purposes, we have chosen to focus on exploring bracketing cases of qualitative ingredients, rather than fine-tuning the model.

Since the clustering signal is strongly influenced by differences between centrals and satellites, then the satellite fraction $F_{\rm sat}(\mstar)$ plays an important role in $\rhalf-$dependent clustering. For {\em fixed} scaling relations $\median{\rhalf^{\rm cens}}{\mhalo}$ and $\median{\rhalf^{\rm sats}}{\mhalo},$ models with different satellite fractions will exhibit distinct $\rhalf-$clustering ratios because $F_{\rm sat}(\mstar)$ controls the relative weighting of the two populations. An explicit demonstration of this point appears in the Appendix, which repeats the analyses in the main body of the text, but for a subhalo catalog with no orphan correction.

@@

Figure \ref{fig:censatsizeratios} shows that the satellites in $\mhalo\approx10^{12}\msun$ halos can be as large as $\lesssim40-50\%$ smaller than their central galaxy counterparts. Since the satellite fraction varies quite significantly with morphology and/or color, we have demonstrated that it is entirely plausible that most of the observed environmental trends of galaxy size can be understood simply in terms of central vs. satellite size differences. The latter reflects not evolution of galaxies inside host halos, but is due to the fact that satellites have reached their peak mass at higher redshifts than the centrals. While it remains to be seen whether this explanation ends up comprising the whole story, one thing is now clear: it is not possible to reliably analyze environmental trends of galaxy $\rhalf$ without properly accounting for satellites.
